Entropy-SGD: Biasing Gradient Descent Into Wide Valleys

### 2017-01-09 doc/chaudhari.choromanska.ea.iclr17.tex
commit before surgery

\author{Pratik Chaudhari$^{1}$, Anna Choromanska$^{2}$, Stefano Soatto$^{1}$, Yann LeCun$^{2,3}$, \red{Carlo Baldassi$^{4}$},\\[0.03in]
\red{\textbf{Christian Borgs$^{5}$, Jennifer Chayes$^{5}$, 
Riccardo Zecchina$^{4}$}}\\[0.05in]
$^{1}$ Computer Science, University of California, Los Angeles.\\
$^{2}$ Courant Institute of Mathematical Sciences, New York University.\\
$^{3}$ Facebook AI Research, New York.\\
$^{4}$ Politecnico di Torino.\\
$^{5}$ Microsoft Research New England, Cambridge.\\ [0.05in]
{\footnotesize
Email:\ \href{mailto:pratikac@ucla.edu}{pratikac@ucla.edu},
\href{mailto:achoroma@cims.nyu.edu}{achoroma@cims.nyu.edu},
\href{mailto:soatto@cs.ucla.edu}{soatto@cs.ucla.edu},
\href{mailto:yann@cs.nyu.edu}{yann@cs.nyu.edu},
\href{mailto:carlobaldassi@gmail.com}{carlobaldassi@gmail.com}}\\[0.03in]
{\footnotesize
\hspace{0.33in} \href{mailto:borgs@microsoft.com}{borgs@microsoft.com},
\href{mailto:jchayes@microsoft.com}{jchayes@microsoft.com},
\red{\href{mailto:leventsagun@gmail.com}{leventsagun@gmail.com},}
\href{mailto:riccardo.zecchina@polito.it}{riccardo.zecchina@polito.it}
}}

@@

\begin{abstract}
This paper proposes a new optimization algorithm called $\entropysgd$ for training deep neural networks that is motivated by the local geometry of the energy landscape. We construct a local-{\color{red}free-}entropy-based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys. Our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of the local-{\color{red}free-}entropy at each update of the weights. We prove that incorporating the local entropy into the objective function results in a smoother energy landscape, and we use uniform stability to show improved generalization bounds over SGD. Our experiments on competitive baselines demonstrate that $\entropysgd$ alone leads to generalization performances that compare with state of the art techniques and has the potential to accelerate training.
\pc{}{We also perform a direct comparison between the information provided by the local entropy measure and the analysis of the Hessian in qualitatively different minima.}
\end{abstract}

>>>>

\author{Pratik Chaudhari$^{1}$, Anna Choromanska$^{2}$, Stefano Soatto$^{1}$, Yann LeCun$^{2,3}$, Carlo Baldassi$^{4}$,\\[0.03in]
\textbf{Christian Borgs$^{5}$, Jennifer Chayes$^{5}$, 
Riccardo Zecchina$^{4}$}\\[0.05in]
$^{1}$ Computer Science, University of California, Los Angeles.\\
$^{2}$ Courant Institute of Mathematical Sciences, New York University.\\
$^{3}$ Facebook AI Research, New York.\\
$^{4}$ Politecnico di Torino.\\
$^{5}$ Microsoft Research New England, Cambridge.\\ [0.05in]
{\footnotesize
Email:\ \href{mailto:pratikac@ucla.edu}{pratikac@ucla.edu},
\href{mailto:achoroma@cims.nyu.edu}{achoroma@cims.nyu.edu},
\href{mailto:soatto@cs.ucla.edu}{soatto@cs.ucla.edu},
\href{mailto:yann@cs.nyu.edu}{yann@cs.nyu.edu},
\href{mailto:carlobaldassi@gmail.com}{carlobaldassi@gmail.com}}\\[0.03in]
{\footnotesize
\hspace{0.33in} \href{mailto:borgs@microsoft.com}{borgs@microsoft.com},
\href{mailto:jchayes@microsoft.com}{jchayes@microsoft.com},
\href{mailto:riccardo.zecchina@polito.it}{riccardo.zecchina@polito.it}
}}

@@

\begin{abstract}
This paper proposes a new optimization algorithm called $\entropysgd$ for training deep neural networks that is motivated by the local geometry of the energy landscape. We construct a local-entropy-based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys. Our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of the local-entropy at each update of the weights. We prove that incorporating the local entropy into the objective function results in a smoother energy landscape, and we use uniform stability to show improved generalization bounds over SGD. Our experiments on competitive baselines demonstrate that $\entropysgd$ alone leads to generalization performances that compare with state of the art techniques and has the potential to accelerate training.
\pc{}{We also perform a direct comparison between the information provided by the local entropy measure and the analysis of the Hessian in qualitatively different minima.}
\end{abstract}
